---
output:
  md_document:
    variant: markdown_github
---

```{r}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
rm(list = ls()) # Clean your environment:

required_packs <- c("tidyverse", "readxl", "xgboost", "pROC", "caret", "dplyr", "ggplot2", "reshape2", "readr")

for (pkg in required_packs) {
    if(!(pkg %in% installed.packages())) install.packages(pkg)
    if(!(pkg %in% (.packages()))) library(pkg, character.only = TRUE) 
}



list.files('./code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
```


## Prep
```{r dataimport}

data <- read_delim("./data/merger_database.csv", delim = ";")

# Use the duplicated() function to identify duplicate rows
duplicate_rows <- duplicated(data)

# Subset the dataset to keep only unique rows
data <- data[!duplicate_rows, ]


# Check for NA values
na_values <- sum(is.na(data))

# Print the number of NA values
print(na_values)

```
```{r exploratoryanalysis}

g <- distdecision(data)

print (g)

```



```{r featureengineering}
#recoding data and generating new dataset

    data <- data %>%
        mutate(Approval_Status = if_else(Approval_Status == "Approved", 1, 0)) %>%
        mutate(Size = case_when(
            Size == "S" ~ 1,
            Size == "I" ~ 2,
            Size == "L" ~ 3,
            TRUE ~ NA_integer_
        )) %>%

        mutate(Joint_Venture = if_else(is.na(Joint_Venture), 0, if_else(Joint_Venture == "Yes", 1, 0))) %>%

        mutate(Joint_Control = if_else(is.na(Joint_Control), 0, if_else(Joint_Control == "Yes", 1, 0))) %>%

        mutate(Barriers_Assess = if_else(is.na(Barriers_Assess), 0, if_else(Barriers_Assess == "Yes", 1, 0))) %>%

        mutate(Barriers_Significant = if_else(is.na(Barriers_Significant), 0, if_else(Barriers_Significant == "Yes", 1, 0))) %>%

        mutate(Entry_Assess = if_else(is.na(Entry_Assess), 0, if_else(Entry_Assess == "Yes", 1, 0))) %>%

        mutate(Expansion_Assess = if_else(is.na(Expansion_Assess), 0, if_else(Expansion_Assess == "Yes", 1, 0))) %>%

        mutate(Countervail_Assess = if_else(is.na(Countervail_Assess), 0, if_else(Countervail_Assess == "Yes", 1, 0))) %>%

        mutate(Efficiency_prov = if_else(is.na(Efficiency_prov), 0, if_else(Efficiency_prov == "Yes", 1, 0))) %>%

        mutate(Counterfactual_Assess = if_else(is.na(Counterfactual_Assess), 0, if_else(Counterfactual_Assess == "Yes", 1, 0))) %>%

        mutate(PublicInterest_Assess = if_else(is.na(PublicInterest_Assess), 0, if_else(PublicInterest_Assess == "Yes", 1, 0))) %>%

        mutate(Overlap_Type = case_when(
            Overlap_Type == "Horizontal" ~ 1,
            Overlap_Type == "Vertical" ~ 2,
            Overlap_Type == "Horizontal and vertical" ~ 3,
            Overlap_Type == "None" ~ 4,
            TRUE ~ 0
        )) %>%

        mutate(Countervail_Finding = if_else(Countervail_Finding %in% c("Yes", "No"), if_else(Countervail_Finding == "Yes", 1, 0), 0)) %>%

        mutate(Efficiency_substantive = if_else(Efficiency_substantive %in% c("Yes", "No"), if_else(Efficiency_substantive == "Yes", 1, 0), 0)) %>%
    
        mutate(Import_Comp = if_else(Import_Comp %in% c("Yes", "No"), if_else(Import_Comp == "Yes", 1, 0), 0)) %>%
    
        mutate(Import_Constraint = if_else(Import_Constraint %in% c("Yes", "No"), if_else(Import_Constraint == "Yes", 1, 0), 0)) %>%

        mutate(Conditions = case_when(
            Conditions %in% c("public interest", "Public Interest") ~ 1,
            Conditions %in% c("competition", "Competition") ~ 2,
            Conditions %in% c("both", "Both") ~ 3,
            is.na(Conditions) ~ 0,
            TRUE ~ NA_integer_
        ))

# Create the merger_df dataframe with selected variables

merger_df <- data %>% 
  select(Approval_Status, Size, Joint_Venture, Joint_Control, Overlap_Type,
         Barriers_Assess, Barriers_Significant,
         Entry_Assess, Expansion_Assess, Countervail_Assess, Countervail_Finding,
         Efficiency_prov, Efficiency_substantive, Counterfactual_Assess,
         PublicInterest_Assess, Year, Phase, Import_Comp, Import_Constraint)

# Print the merger_df dataframe
summary(merger_df)

```

```{r exploratoryanalysis2}

g <- heatmap(merger_df)
print(g)

# # Create the boxplot
# ggplot(merger_df, aes(x = Phase, y = Approval_Status, group = Phase)) +
#     geom_boxplot(fill = "lightblue", color = "black") +
#     labs(x = "Phase", y = "Approval Status", title = "Boxplot of Phase by Approval Status")
# 
# ggplot(merger_df, aes(x = Phase, y = Approval_Status, group = Phase)) +
#     geom_boxplot(fill = "lightblue", color = "black") +
#     labs(x = "Phase", y = "Approval Status", title = "Boxplot of Phase by Approval Status")

```



## Baseline Accuracy 
This is ZeroR baseline accuracy prediction
```{r baselineacc}

g <- baseline_acc(merger_df)

print(g)

```





## XGBoost
```{r prep}
# Split your data into training and testing sets (e.g., 80% for training and 20% for testing)
set.seed(123)
train_indices <- sample(1:nrow(merger_df), 0.8 * nrow(merger_df))
train_data <- merger_df[train_indices, ]
test_data <- merger_df[-train_indices, ]
```


```{r xgboost}
# Define the target variable
target = 'Approval_Status'

class(train_data[[target]])

#generating output vector
output_vector = merger_df[, target] == 1


# Create the XGBoost matrix (DMatrix) for training and testing
dtrain <- xgb.DMatrix(data = as.matrix(train_data[, !colnames(train_data) %in% target]), label = as.numeric(train_data[[target]]))
dtest <- xgb.DMatrix(data = as.matrix(test_data[, !colnames(test_data) %in% target]), label = as.numeric(test_data[[target]]))


# Train the XGBoost model
model <- xgboost(data = dtrain, subsample = 0.5, nrounds = 10, objective = "binary:logistic")
```

```{r prediction}

pred <- predict(model, dtest)

err <- mean(as.numeric(pred > 0.5) != (test_data[[target]]))
print(paste("test-error=", err))

# Evaluate the model's performance (e.g., accuracy, confusion matrix, etc.)
# Example: Calculate accuracy
# actual_labels <- test_data[, target]
# accuracy <- sum(predictions > 0.5 == actual_labels) / length(actual_labels)

```
## Benchmark model fitting (logistic regression)
```{r benchmark}
lgt_model <- glm(Approval_Status ~. - Approval_Status, data = train_data, family = "binomial")

lgt_preds <- predict(lgt_model, newdata = test_data, type = "response")
table(lgt_preds >= 0.5, target == 1)

print(paste("Logistic regression score:", auc(target, lgt_preds)))
```



```{r visualinterp} 
#visual interpretation of initial results

g <- importance_matrix(merger_df)

print(g)


```
Finding the variable importance in the model and subsetting the variable list:

As it stands, it appears as though all the variables have at least a small effect on the outcome of Approval Status of mergers when applied to the Competition Commission's merger database. 

If anything, we can attempt to remove 'Joint_Control' and 'Countervail_Assess' and see if this improves model accuracy. 

- we need to run with standard hyperparameters, adjust and rerun, adjust and rerun.... 
```{r chisquaretests}

g <- chisquaretests(merger_df)

print(g)
```



